Interoception Scale - Data Cleaning

### 2024-11-28 study1/analysis/1_cleaning.qmd
Changes to cleaning files

#df <- read_csv("C:/Users/emmal/OneDrive - University of Sussex/GitHub/Placement/InteroceptionScale/study1/data/rawdata_participants.csv")

df <- read.csv("https://raw.githubusercontent.com/RealityBending/InteroceptionScale/refs/heads/main/study1/data/rawdata_participants.csv")
```

@@

p_age <- estimate_density(df$Age) |>
  normalize(select = y) |>
  mutate(y = y * 86) |>  # To match the binwidth
  ggplot(aes(x = x)) +
  geom_histogram(data=df, aes(x = Age, fill=Gender), bins=28) +
  # geom_line(aes(y = y), color = "orange", linewidth=2) +
  geom_vline(xintercept = mean(df$Age), color = "#ca0020", linewidth=1.5) +
  # geom_label(data = data.frame(x = mean(df$Age) * 1.15, y = 0.95 * 75), aes(y = y), color = "red", label = paste0("Mean = ", format_value(mean(df$Age)))) +
  scale_fill_manual(values = c("Male"= "#64B", "Female"= "#8EB69B", "Other"="orange")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Age and Gender", y = "Number of Participants", color = NULL, subtitle = "Distribution of participants' age and gender") +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )
p_age
```

@@

# did not add education disiplines

p_edu <- df |>
  mutate(Education = fct_relevel(Education, "Secondary (technical) School", "Elementary School", "High School", "Bachelor", "Master", "Doctorate")) |>
  ggplot(aes(x = Education)) +
  geom_bar(aes(fill = Education)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Education", y = "Number of Participants", subtitle = "Participants per achieved education level") +
  theme_modern(axis.title.space = 15) +
  # scale_fill_manual(values = c("Secondary (technical) school"="#0571B0","High school"= "#d01c8b", "Bachelor"="#FDB863", "Master"="#E66101", "Doctorate" =  "#ca0020")) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.5), angle = 45, hjust =1),
    axis.title.x = element_blank()
  )

@@

# Student

table(df$Student)
```

## Ethnicity

# Ethnicity

p_eth <- df |>
  mutate(Ethnicity = fct_relevel(Ethnicity,"East Asian", "Mixed", "White", "Black", "persian", "South Asian", "Hispanic/Latino")) |>
  ggplot(aes(x = Ethnicity)) +
  geom_bar(aes(fill = Ethnicity)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Ethnicity", y = "Number of Participants", subtitle = "Participants per Ethnicity") +
  theme_modern(axis.title.space = 15) +
  #scale_fill_manual(values = c("East Asian"="#0571B0","Mixed"= "#d01c8b", "White"="#cc66cc", "Black"="#E66101", "persian" =  "#ca0020", "South Asian"="#5E3C99", "Hispanic/Latino"="#92C5DE")) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.6), angle = 45, hjust = 1),
    axis.title.x = element_blank())

@@

## Residence

@@

# Country participants are currently living

p_cou <- df |>
  mutate(Country = fct_relevel(Country,"United Kingdom", "Ireland", "USA", "Australia")) |>
  ggplot(aes(x = Country)) +
  geom_bar(aes(fill = Country)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Country of Residence", y = "Number of Participants", subtitle = "Participants per Country they are currently living") +
  theme_modern(axis.title.space = 15) +
  #scale_fill_manual(values = c("United Kingdom"="#ca0020","Ireland"= "#d01c8b", "United States"="#FDB863", "Australia"="#E66101"))+
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.7),  angle = 45, hjust =1),
    axis.title.x = element_blank())

p_cou

p_country <- df |>
  mutate(Country = case_when(
    Country=="United States"~ "USA",
    Country=="United Kingdom" ~ "UK",
    TRUE ~ Country
  ))|>
  dplyr::select(region = Country) |>
  group_by(region) |>
  summarize(n = n()) |>
  right_join(map_data("world"), by = "region") |>
  # mutate(n = replace_na(n, 0)) |>
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = n)) +
  scale_fill_gradientn(colors = c("#E66101", "#ca0020", "#cc66cc")) +
  labs(fill = "N") +
  theme_void() +
  labs(title = "Country of Residence", subtitle = "Number of participants by country")  +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2))
  )

p_country

p_feed <- estimate_density(df$Experiment_Enjoyment) |>
  normalize(select = y) |>
  mutate(y = y * 86) |>  # To match the binwidth
  ggplot(aes(x = x)) +
  geom_bar(data=df, aes(x = Experiment_Enjoyment, fill = Condition), bins=28) +
  scale_fill_manual(values = c("Dimensions"= "#E66101","Domains" = "#ca0020", "Random"="#d01c8b")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Experiment Enjoyment and Condition", y = "Number of Participants", fill = " ", x = "Ratings of enjoyment on a scale 1-5", subtitle = "Distribution of Participants' Experiment Enjoyment and Condition") +
  # geom_label(data = data.frame(x = mean(df$Experiment_Enjoyment) * 1.15, y = 0.95 * 75), aes(y = y), color = "red", label = paste0("Mean = ", format_value(mean(df$Experiment_Enjoyment)))) +
  facet_wrap(~Condition) +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    #axis.title.x = element_blank()
  )
p_feed

  #geom_line(aes(y = y), color = "orange", linewidth=2) +
#geom_vline(xintercept = mean(df$Experiment_Enjoyment), color = "#ca0020", linewidth=1.5) +
```

>>>>

# df <- read_csv("C:/Users/emmal/OneDrive - University of Sussex/GitHub/Placement/InteroceptionScale/study1/data/rawdata_participants.csv")

df <- read.csv("https://raw.githubusercontent.com/RealityBending/InteroceptionScale/refs/heads/main/study1/data/rawdata_participants.csv")

## Attention Checks

df <- df |>
  dplyr::mutate(
    A1 = dplyr::case_when(
      # "I always know that I am attentively doing a study"
      Sexual_State_A == 6 | Sexual_State_A == 5 ~ "Pass",
      TRUE ~ "Fail"),
    A2 = dplyr::case_when(
      # "Even if I am anxious, I should now answer all the way to the left"
      Anxious_SkinThermo_A == 0 ~ "Pass",
      TRUE ~ "Fail"),
    A3 = dplyr::case_when(
      # "I often experience sensations, and I will answer zero to this question"
      Nociception_ColonBladder_A == 0 ~ "Pass",
      TRUE ~ "Fail"),
    A4 = dplyr::case_when(
      # "In general, I am very sensitive and attentive to the questions I am       currently answering"
      Sensitivity_Cardiac_A == 6 | Sensitivity_Cardiac_A == 5 ~ "Pass",
      TRUE ~ "Fail"),
    A5 = dplyr::case_when(
      # "I often pay attention to the answers I am giving"
      Sensitivity_Gastric_A == 6 | Sensitivity_Gastric_A == 5 ~ "Pass",
      TRUE ~ "Fail"),
    A6 = dplyr::case_when(
      # "I can always accurately answer to the left on this question to show that I am reading it" 
      Accuracy_Respiratory_A == 0 | Accuracy_Respiratory_A == 1 ~ "Pass",
      TRUE ~ "Fail"),
    A7 = dplyr::case_when(
      # "I can always accurately perceive that to this question I should answer the lowest option"
      Accuracy_Genital_A == 0 ~ "Pass",
      TRUE ~ "Fail"),
    A8 = dplyr::case_when(
      #  "Sometimes I notice that I need to answer all the way to the right"
      Confusion_ColonBladder_A == 6 ~ "Pass",
      TRUE ~ "Fail"),
  )

# Create colum with the total number of Failed attention checks 
df$Total_Attention_Failed <- apply(df[, c("A1", "A2", "A3", "A4", "A5", "A6", "A7", "A8")], 1, function(x) sum((x == "Fail")))

p_fail <- df |>
  ggplot(aes(x = Total_Attention_Failed)) +
  geom_bar(aes(fill = factor(Total_Attention_Failed))) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Failed Attention Checks", y = "Number of Participants", subtitle = "Number of failed attention checks per participant") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.5)),
    axis.title.x = element_blank()
  )

p_fail

```{r}

removed_participants <- nrow(df |> filter(Total_Attention_Failed >= 1))

df <- df |>
  filter(Total_Attention_Failed == 0)
```

`r removed_participants` participants were removed for having failed at least 1 attention check.

p_feed <- estimate_density(df$Experiment_Enjoyment) |>
  normalize(select = y) |>
  mutate(y = y * 86) |>  # To match the binwidth
  ggplot(aes(x = x)) +
  geom_bar(data=df, aes(x = Experiment_Enjoyment, fill = Condition), bins=28) +
  scale_fill_manual(values = c("Dimensions"= "#ff944d","Domains" = "#ca0020", "Random"="#ff80d5")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Experiment Enjoyment and Condition", y = "Number of Participants", fill = " ", x = "Ratings of enjoyment on a scale 1-5", subtitle = "Distribution of Participants' Experiment Enjoyment and Condition") +
  facet_wrap(~Condition) +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
  )
p_feed

## Final Sample

::: panel-tabset

## Summary

-The final sample includes `r report::report_participants(df, age = "Age", sex = NA, gender = "Gender")`.

@@

```{r p_demographics_age_gender}
#| warning: false
#| code-fold: true
#| error: false

p_age <- estimate_density(df$Age) |>
  normalize(select = y) |>
  mutate(y = y * 86) |>  # To match the binwidth
  ggplot(aes(x = x)) +
  geom_histogram(data=df, aes(x = Age, fill=Gender), bins=28) +
  geom_vline(xintercept = mean(df$Age), color = "#ca0020", linewidth=1.5) +
  scale_fill_manual(values = c("Male"= "#64B", "Female"= "#8EB69B", "Other"="orange")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Age and Gender", y = "Number of Participants", color = NULL, subtitle = "Distribution of participants' age and gender") +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),)

p_age
```

@@

```{r p_demographics_education}
#| warning: false
#| code-fold: true
#| error: false

# Did not add education disciplines

p_edu <- df |>
  ggplot(aes(x = Education)) +
  geom_bar(aes(fill = Education)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Education", y = "Number of Participants", subtitle = "Participants per achieved education level") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.5), angle = 45, hjust =1),
    axis.title.x = element_blank()
  )

@@

# Student
table(df$Student) |>
  as.data.frame() |>
  rename(Student = Var1, Frequency = Freq) |>
  knitr::kable()
```

## Ethinicity

```{r p_demographics_ethnicity}
#| warning: false
#| code-fold: true
#| error: false

p_eth <- df |>
  ggplot(aes(x = Ethnicity)) +
  geom_bar(aes(fill = Ethnicity)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Ethnicity", y = "Number of Participants", subtitle = "Participants per Ethnicity") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.6), angle = 45, hjust = 1),
    axis.title.x = element_blank())

@@

## Country of Residence

```{r p_demographics_country}
#| warning: false
#| code-fold: true
#| error: false

p_cou <- df |>
 ggplot(aes(x = Country)) +
 geom_bar(aes(fill = Country)) +
 scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
 scale_fill_viridis_d(guide = "none") +
 labs(title = "Country of Residence", y = "Number of Participants", subtitle = "Participants per Country they are currently living") +
 theme_modern(axis.title.space = 15) +
 theme(
   plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
   plot.subtitle = element_text(size = rel(1.2), vjust = 7),
   axis.text.y = element_text(size = rel(1.1)),
   axis.text.x = element_text(size = rel(0.7), angle = 45, hjust = 1),
   axis.title.x = element_blank())

p_cou
```

```{r p_demographics_map}
#| warning: false
#| code-fold: true
#| error: false

p_map <- df |>
  mutate(Country = case_when(
    Country=="United States"~ "USA",
    Country=="United Kingdom" ~ "UK",
    TRUE ~ Country
  ))|>
  dplyr::select(region = Country) |>
  group_by(region) |>
  summarize(n = n()) |>
  right_join(map_data("world"), by = "region") |>
  # mutate(n = replace_na(n, 0)) |>
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = n)) +
  scale_fill_gradientn(colors = c("#E66101", "#ca0020", "#cc66cc")) +
  labs(fill = "N") +
  theme_void() +
  labs(title = "Country of Residence", subtitle = "Number of participants by country")  +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2))
  )

p_map
```
:::

```{r p_summary}
#| warning: false
#| code-fold: true
#| error: false

p_map 
p_age 
p_edu
```

## Save

@@

write.csv(df, "../data/data_participants.csv", row.names = FALSE)
```
